% =====================================================================
%  FWM_physics.tex
%  A physics-first walkthrough of the GABES four-wave-mixing engine.
%
%  Personal study notes: every physical step from the atomic level
%  scheme to the squeezing / biphoton observables, mapped onto the
%  actual code path (gabes/schemes/fwm.py, core.py, observables.py,
%  doppler.py, atoms.py, constants.py, hyperfine.py, species.py).
%
%  Compile with pdflatex (no external packages beyond a standard TeX
%  Live / MiKTeX install; no images, no shell-escape).
% =====================================================================
\documentclass[11pt]{article}

\usepackage[a4paper,margin=2.4cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{verbatim}
\usepackage[colorlinks=true,linkcolor=blue!50!black,urlcolor=blue!50!black,
            citecolor=blue!50!black]{hyperref}

% ---- light-weight macros (avoid depending on physics/braket pkgs) ----
\newcommand{\ket}[1]{\lvert #1 \rangle}
\newcommand{\bra}[1]{\langle #1 \rvert}
\newcommand{\braket}[2]{\langle #1 \vert #2 \rangle}
\newcommand{\ketbra}[2]{\lvert #1 \rangle\!\langle #2 \rvert}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\vecop}{\operatorname{vec}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\rd}{\,\mathrm{d}}
\newcommand{\eff}{\mathrm{eff}}
\newcommand{\HF}{\mathrm{HF}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\Lcal}{\mathcal{L}}
\newcommand{\Dcal}{\mathcal{D}}
\newcommand{\rhobar}{\bar{\chi}}
\DeclareMathOperator{\Real}{Re}
\DeclareMathOperator{\Imag}{Im}

\newtheorem{remark}{Remark}

\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}

\title{\textbf{Four-Wave Mixing in Hot \boldmath$^{85}$Rb:\\
A Physics-First Walkthrough of the GABES Engine}\\[0.3em]
\large Current FWM Physics Model --- revised 2026-07-03\\
(first-principles coupling, collisional decoherence, corrected IDS formula,\\
hardened Ultra noise, predictive biphoton)}
\author{Personal study notes}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
\noindent
These notes derive, step by step, every piece of physics that the GABES
four-wave-mixing (FWM) code evaluates --- from the open-system master equation of
the $^{85}$Rb $D_1$ double-$\Lambda$ system, through the Doppler average and the
truncated-Floquet sideband steady state, to the Maxwell--Bloch propagation that
turns the atomic susceptibility into seed/conjugate gain, intensity-difference
squeezing, and (in the second mode) a spontaneous-FWM biphoton waveform.
The macroscopic coupling is now first-principles: the hyperfine Clebsch--Gordan
structure inside $\rhobar$ is completed by the explicit population/degeneracy
normalization $p_F/[2(2I+1)]$ and the pure-$^{85}$Rb CRC vapour density, leaving
only a dimensionless residual line-strength calibration ($\approx1$). The
dissipator carries density- and temperature-dependent collisional decoherence
(optical self-broadening $\Gamma+\beta N$ and ground-state Raman collisions), so
the achievable squeezing peaks near $121$--$131\,^\circ$C and degrades at higher
temperature, as observed. The ideal intensity-difference noise uses the
conserved-difference Bogoliubov result
$S_{\text{ideal}}=(G_s-G_c)^2/(G_s+G_c)$ [$=1/(2G_s-1)$ in the lossless limit],
replacing an older expression that overestimated squeezing at high gain. Seeded
squeezing includes the reference $0.32^\circ$ pump--probe longitudinal mismatch
from Balanced fidelity up; High fidelity adds a conservative $\chi$-reused
refractive correction and segmented propagation; Ultra adds 64-segment dynamic
pump-budget depletion, Gaussian beam overlap, in-cell loss/noise, Zeeman
diagnostics, and --- intrinsic since the v5 hardening --- an excess-noise model
with arm-resolved linear absorption of probe and conjugate plus OD-anchored pump
scatter. Biphoton mode defaults to a predictive first-principles Doppler-averaged
waveform (Autler--Townes two-photon denominator, velocity-class coherent sum on
an auto-refined grid, natural-linewidth decay) and uses vector phase matching:
calibrated longitudinal $\Delta k_z$ plus absolute transverse $\Delta k_x$, with
a two-dimensional signal--idler vector acceptance map in Fine mode. The heavy OBE
grid can be evaluated by optional compiled Numba kernels; the $(\Delta,T)$
squeezing scan builds one master $\rhobar(\delta,\Delta_\eff)$ table \emph{per
temperature} (the collisional dissipator makes $\rhobar$ $T$-dependent). The
emphasis is on the \emph{physics}; the code is referenced only as a map so that
each equation can be located in \code{gabes/}. The direct
\code{compute\_spectrum} path is regression-anchored to
G.~Sim, H.~Kim, and H.~S.~Moon, \emph{Sci.\ Rep.}\ \textbf{15}, 7727 (2025).
\end{abstract}

\tableofcontents
\bigskip\hrule

% =====================================================================
\section{Purpose and the object we compute}
% =====================================================================

Non-degenerate four-wave mixing in a hot alkali vapour is a third-order parametric
process. A strong \emph{pump} drives a multilevel atom; a weak \emph{seed} (probe)
injected on one Raman leg stimulates the emission of a \emph{conjugate} on the
other leg. Energy and momentum conservation tie the two generated photons
together, so signal and conjugate emerge as a two-mode squeezed (twin-beam) state:
their \emph{intensity difference} is quieter than the shot-noise limit, even though
each beam alone is bright and thermal-like.

The code answers two questions:

\begin{enumerate}[leftmargin=2.2em]
  \item \textbf{Seeded / squeezing mode.} Given a one-photon detuning $\Delta$,
  cell temperature $T$, pump power, seed power, and detection efficiency, what is
  the seed power gain $G_s$, the conjugate gain $G_c$, and the intensity-difference
  squeezing $S$ (in dB) as a function of the two-photon detuning $\delta$? This is
  the physics of Sim \emph{et al.}, and is the regression-anchored core.
  \item \textbf{Spontaneous / biphoton mode.} For a cascade or diamond topology
  with \emph{no} seed, what is the time-correlation $g^{(2)}_{SI}(\tau)$ of the
  spontaneously generated signal--idler pair, its width, pair rate, and
  coincidence-to-accidental ratio (CAR)? The default Predictive model computes
  the waveform, width and bandwidth from first principles, with a
  reference-anchored absolute rate; a legacy Calibrated model is kept for
  comparison. Either way it is a source estimate, not a full quantum-Langevin
  propagation.
\end{enumerate}

The whole numerical machine exists to evaluate one object accurately: the
\emph{atomic susceptibility matrix} $\rhobar(\delta,\Delta;v,T)$ that links the
induced optical polarizations on the two Raman legs to the seed and conjugate
fields. Everything before it (Hamiltonian, dissipator, Floquet, Doppler) builds
that matrix; everything after it (Maxwell--Bloch, saturation, squeezing formula)
reads it out.

\paragraph{Code map.} The seeded pipeline lives in
\code{gabes/schemes/fwm.py::compute\_spectrum}; the physics-agnostic OBE engine is
\code{gabes/core.py}; the susceptibility$\,\to\,$observable layer is
\code{gabes/observables.py}; the level model and rates are
\code{gabes/atoms.py} and \code{gabes/hyperfine.py}; the thermal average is
\code{gabes/doppler.py}; physical constants and field helpers are
\code{gabes/constants.py}. The biphoton mode is the
\code{GenericFWMSolver} class in the same \code{fwm.py}.

% =====================================================================
\section{The double-\boldmath$\Lambda$ system and its energy scales}
% =====================================================================

\subsection{Levels}

The seeded model uses the $^{85}$Rb $D_1$ line ($5S_{1/2}\!\to\!5P_{1/2}$,
$\lambda \approx 795\,$nm) reduced to four levels (\code{atoms.double\_lambda\_rb85};
the builder now takes the dephasing rates $\gamma_{gg}$ and
$\gamma_{\text{opt}}$ as arguments, see Sec.~\ref{sec:collisional}):

\begin{center}
\begin{tabular}{clcl}
\toprule
index & state & manifold & role \\
\midrule
$g_1$ & $\ket{5S_{1/2},F=2}$ & ground & lower Raman state \\
$g_2$ & $\ket{5S_{1/2},F=3}$ & ground & upper Raman state \\
$e_2$ & $\ket{5P_{1/2},F'=2}$ & excited & optical \\
$e_3$ & $\ket{5P_{1/2},F'=3}$ & excited & optical \\
\bottomrule
\end{tabular}
\end{center}

The two ground hyperfine states are split by
$\nu_{\HF}=\nu_{\text{ground HF}}=3.0357\,$GHz; the two excited states by
$\nu^{(e)}_{\HF}=361.58\,$MHz. The geometry is ``double-$\Lambda$'': each ground
state couples optically to \emph{both} excited states, and the two ground states
share the excited manifold. That shared manifold is what lets a seed on the
$g_2\!\leftrightarrow\!e$ leg transfer coherence to the $g_1\!\leftrightarrow\!e$
leg and generate the conjugate --- the four ``waves'' are pump, pump, seed,
conjugate.

\subsection{Hyperfine line strengths (Clebsch--Gordan)}

Each $F\!\to\!F'$ link carries a relative strength $C_F^2$ from the Wigner
$3j/6j$ sum (\code{hyperfine.CF2}), exact rationals for $^{85}$Rb $D_1$:
\begin{equation}
C_F^2(2,2)=\tfrac{10}{81},\quad
C_F^2(2,3)=\tfrac{35}{81},\quad
C_F^2(3,2)=\tfrac{35}{81},\quad
C_F^2(3,3)=\tfrac{28}{81}.
\label{eq:cf2}
\end{equation}
These factors enter the physics in three distinct places, and using one consistent
quantity for all three is what makes the model self-consistent:
\begin{enumerate}[leftmargin=2.2em]
  \item \textbf{Coherent drive.} The effective Rabi frequency on link $g\!\to\!e$
  is scaled by $\sqrt{3\,C_F^2(F_g,F_{e})}$ (\code{TRANSITION\_DIPOLE\_SCALE});
  the factor 3 is the $J\!\to\!J'$ reduction of the $D_1$ reduced matrix element.
  \item \textbf{Polarization readout.} The macroscopic polarization on a leg is the
  same CG-weighted sum of optical coherences (Sec.~\ref{sec:chi}).
  \item \textbf{Spontaneous emission branching.} An excited state $e$ decays to
  ground $g$ at the branched rate
  \begin{equation}
  \Gamma_{e\to g}=\Gamma\,\frac{C_F^2(F_g,F_{e})}{\sum_{g'}C_F^2(F_{g'},F_{e})},
  \qquad \sum_g \Gamma_{e\to g}=\Gamma,
  \label{eq:branching}
  \end{equation}
  so total population leaves each excited state at the natural rate $\Gamma$
  (\code{atoms.double\_lambda\_rb85}).
\end{enumerate}

\subsection{Energy scales and why the rotating frame is mandatory}

The relevant frequencies span fifteen orders of magnitude: the optical carrier is
$\omega_{D_1}/2\pi=377\,$THz, the hyperfine splitting $3\,$GHz, the natural
linewidth $\Gamma/2\pi=5.746\,$MHz, the ground-coherence decay $100\,$kHz. No
solver integrates the bare optical oscillation. Instead we move to a frame
co-rotating with the pump and apply the rotating-wave approximation (RWA): all
optical carriers are removed, leaving \emph{detunings} of order MHz--GHz on the
diagonal and slowly varying Rabi couplings off-diagonal. Throughout, energies are
expressed as angular frequencies (i.e.\ $\hbar\!\equiv\!1$ in the Hamiltonian), so
``$H$'' below is really $H/\hbar$ in rad\,s$^{-1}$.

% =====================================================================
\section{Units, fields, and the vapour density}
% =====================================================================

\subsection{Rabi frequency from beam power}

A beam of power $P$ and $1/e^2$ intensity radius $w_0$ has peak intensity
$I=2P/(\pi w_0^2)$. The code converts this to a Rabi frequency through the
saturation intensity (\code{constants.rabi\_freq}):
\begin{equation}
I=\frac{2P}{\pi w_0^2},
\qquad
\Omega=\Gamma\sqrt{\frac{I}{2 I_{\mathrm{sat}}}},
\qquad I_{\mathrm{sat}}=4.484\,\mathrm{mW\,cm^{-2}}.
\label{eq:rabi}
\end{equation}
Equation~\eqref{eq:rabi} is just the standard two-level relation
$I/I_{\mathrm{sat}}=2\Omega^2/\Gamma^2$ solved for $\Omega$; tying $\Omega$ to
$I_{\mathrm{sat}}$ keeps the drive strength anchored to a measured quantity rather
than to an absolute dipole. With the paper geometry
($w_{\text{pump}}=530\,\mu$m, $w_{\text{seed}}=330\,\mu$m, both $1/e^2$
\emph{radii}), $600\,$mW of pump gives $\Omega_{\text{pump}}/2\pi$ of order a
GHz, while $8\,\mu$W of seed gives $\Omega_{\text{seed}}/2\pi$ of order MHz --- a
clean separation of ``strong pump, weak seed'' that the linear-response treatment
below relies on.

\subsection{Number density}

The atomic density sets the optical depth and hence the gain. The seeded model
uses the \emph{pure-$^{85}$Rb} CRC vapour-pressure fit
(\code{hyperfine.number\_density}, isotopic fraction $f_{85}=1$) --- the same
density that lets the validated absorption scheme reproduce the lab AutoOD
transmission to $<0.1\,\%$:
\begin{equation}
\log_{10}P_{\text{Pa}} =
\begin{cases}
\;9.863-4215/T, & T<312.45\,\mathrm{K}\ \ (\text{solid Rb}),\\[3pt]
\;9.318-4040/T, & T\ge312.45\,\mathrm{K}\ \ (\text{liquid Rb}),
\end{cases}
\qquad
N=\frac{P}{k_B T}.
\label{eq:crc}
\end{equation}
At the reference $121\,^\circ$C this gives $N=2.15\times10^{19}\,\mathrm{m^{-3}}$.
The enriched cell of the reference experiment is very nearly pure $^{85}$Rb, so
the previously used natural-abundance Steck fit (\code{atoms.rb85\_density},
$f_{85}=0.7217$; retained for the other schemes) understated $N$ for the FWM cell
by roughly the abundance factor. The exponential temperature dependence of $P$ is
the origin of the model's extreme sensitivity to $T$: a few degrees changes $N$
by tens of percent, and the gain is exponential in $N$ (Sec.~\ref{sec:mb}). The
biphoton mode uses the per-isotope elemental densities in
\code{species.number\_density} (CRC fit for Rb, Steck for Cs).

% =====================================================================
\section{Open-system dynamics: the master equation}
% =====================================================================

The atom is open: it spontaneously emits, and ground coherences dephase from
collisions and transit. The state is therefore a density matrix $\rho$ obeying a
Gorini--Kossakowski--Sudarshan--Lindblad (GKSL) master equation,
\begin{equation}
\dot\rho=-\ii\,[H,\rho]+\Dcal[\rho],
\label{eq:gksl}
\end{equation}
with $H$ the rotating-frame Hamiltonian (Sec.~\ref{sec:H}) and $\Dcal$ the
dissipator (Sec.~\ref{sec:diss}). We only ever want the \emph{steady state}
($\dot\rho=0$), since the experiment is continuous-wave.

\subsection{The Hamiltonian}
\label{sec:H}

In the rotating frame the diagonal of $H$ carries detunings and hyperfine
splittings, and the off-diagonal carries the (RWA) drives. The static part at zero
effective detuning is built by
\code{static\_hamiltonian\_at\_Deff\_zero}:
\begin{equation}
H_0=
\underbrace{\delta\,\ketbra{g_2}{g_2}
-\omega^{(e)}_{\HF}\,\ketbra{e_2}{e_2}}_{\text{diagonal: two-photon \& excited HF}}
\;+\;
\underbrace{\sum_{\text{legs}}\frac{\Omega_{ge}}{2}\big(\ketbra{g}{e}+\ketbra{e}{g}\big)}_{\text{RWA drives}},
\label{eq:H0}
\end{equation}
where $\delta$ is the two-photon (Raman) detuning, $\omega^{(e)}_{\HF}=2\pi
\nu^{(e)}_{\HF}$ places $e_3$ at zero and $e_2$ below it, and the leg drive is
$\Omega_{ge}=\Omega\sqrt{3 C_F^2(F_g,F_e)}$ from Eq.~\eqref{eq:cf2}. The one-photon
detuning $\Delta$ and the atomic velocity do \emph{not} appear in $H_0$; they enter
later as a single shift $\Delta_\eff$ on the excited diagonal (Sec.~\ref{sec:dopp}),
which is the trick that makes the Doppler average cheap.

\paragraph{Which field sits where (the branch).} A subtlety unique to FWM: the
pump and the weak fields can be assigned to the two legs in two inequivalent ways,
corresponding to the two Raman resonances. The code labels them
$\text{branch}=\mp1$. For the standard red-detuned seed (branch $-1$), the pump
$\Omega_A$ drives the $g_1$ leg and the seed $\Omega_s$ drives the $g_2$ leg in
$H_0$; for branch $+1$ the assignment swaps. These are \emph{physically distinct
mode pairs} and must be computed separately --- never summed into one
susceptibility matrix (see Sec.~\ref{sec:branch} and the README ``Traps'').

\subsection{The dissipator}
\label{sec:diss}

The dissipator is a sum of GKSL terms, one per irreversible channel,
\begin{equation}
\Dcal[\rho]=\sum_k\Big(L_k\rho L_k^\dagger
   -\tfrac12 L_k^\dagger L_k\,\rho-\tfrac12\rho\,L_k^\dagger L_k\Big).
\label{eq:lindblad}
\end{equation}
Two kinds of channel appear (\code{atoms.AtomModel.\_build\_lindblad}):

\begin{itemize}[leftmargin=1.6em]
  \item \textbf{Spontaneous emission}: jump operators
  $L_{e\to g}=\sqrt{\Gamma_{e\to g}}\,\ketbra{g}{e}$ with the CG-branched rates of
  Eq.~\eqref{eq:branching}. The ``recycling'' term $L\rho L^\dagger$ refeeds the
  emitted population into the ground states; the anticommutator term damps the
  optical coherences at $\Gamma/2$ (the homogeneous optical linewidth).
  \item \textbf{Ground-state (Raman) dephasing}: a decay of the
  $g_1\!\leftrightarrow\!g_2$ coherence,
  $\dot\rho_{g_1g_2}\supset-\gamma_{gg}\rho_{g_1g_2}$. This is the rate that
  sets the width of the Raman/two-photon resonance and ultimately limits the
  squeezing bandwidth. It is no longer a single constant: it is a transit-time /
  residual floor ($\gamma_{\text{floor}}/2\pi=100\,$kHz,
  \code{constants.GAMMA\_GG}) plus a density-dependent Rb--Rb collision term
  (Sec.~\ref{sec:collisional}).
  \item \textbf{Optical collisional dephasing}: an extra decay
  $\gamma_{\text{opt}}$ on every optical $(g,e)$ coherence, representing Rb--Rb
  self-broadening of the optical line on top of the natural $\Gamma/2$
  (Sec.~\ref{sec:collisional}). With $\gamma_{\text{opt}}=0$ (the registry
  default) the natural-linewidth model is recovered --- this is what the pinned
  kernel regressions use.
\end{itemize}

\begin{remark}
The general \code{AtomModel} also supports \emph{polarization-grouped} jump
operators ($\Sigma_q$ form) that transfer coherence during spontaneous emission
(used by the magneto-optics schemes). The four-level FWM model only needs the
population-refeeding form above.
\end{remark}

\subsection{Density-dependent collisional decoherence}
\label{sec:collisional}

Sim \emph{et al.}\ attribute the degradation of squeezing at high vapour
temperature to collision-driven decoherence, and the model now carries that
physics explicitly. Two dephasing channels grow with density; both are built per
temperature by \code{fwm.collisional\_atom(T)}, which calls
\code{atoms.double\_lambda\_rb85(gamma\_gg, gamma\_opt)}:

\begin{enumerate}[leftmargin=2.2em]
  \item \textbf{Optical self-broadening.} Resonant Rb--Rb collisions broaden the
  optical line linearly in density,
  \begin{equation}
  \Gamma_\eff=\Gamma+\beta N,
  \qquad
  \beta/2\pi=0.69\times10^{-7}\,\mathrm{Hz\,cm^{3}},
  \label{eq:selfbroad}
  \end{equation}
  (\code{hyperfine.self\_broadened\_gamma}; the $\beta$ value is the one the
  AutoOD-validated absorption path uses). The natural $\Gamma$ is already carried
  by the spontaneous-emission jump operators, so the dissipator receives only the
  \emph{added} collisional half-width
  \begin{equation}
  \gamma_{\text{opt}}=\tfrac12\big(\Gamma_\eff-\Gamma\big)=\tfrac12\,\beta N
  \label{eq:gammaopt}
  \end{equation}
  as a pure-dephasing rate on each optical $(g,e)$ coherence. At the reference
  $121\,^\circ$C ($N=2.15\times10^{19}\,\mathrm{m^{-3}}$) this is
  $\gamma_{\text{opt}}/2\pi\approx0.74\,$MHz --- already a $\sim\!26\,\%$
  addition to the natural half-width $\Gamma/2$, and it grows exponentially with
  $T$ through $N(T)$.
  \item \textbf{Ground Raman collisions.} The ground-coherence relaxation is
  \begin{equation}
  \gamma_{gg}(T,N)=\gamma_{\text{floor}}+N\,\sigma\,\bar v_{\text{rel}},
  \qquad
  \bar v_{\text{rel}}=\sqrt{\frac{8k_BT}{\pi\,(m/2)}},
  \label{eq:gammagg}
  \end{equation}
  (\code{constants.ground\_coherence\_dephasing}), with the transit/residual
  floor $\gamma_{\text{floor}}/2\pi=100\,$kHz anchored at the validated operating
  point and the Rb--Rb spin-exchange cross-section
  $\sigma=1.9\times10^{-14}\,\mathrm{cm^{2}}$. At $121\,^\circ$C the collisional
  term adds only $\sim\!3\,$kHz --- the transit floor dominates over the whole
  pure-cell range --- but the rate now scales correctly with density.
\end{enumerate}

The physical consequence is a genuine interior optimum in temperature: raising
$T$ raises $N$ and hence the parametric gain exponentially, but the same $N$
degrades the optical and Raman coherences that the gain is built from, so the
achievable squeezing peaks near $121$--$131\,^\circ$C and then \emph{worsens} at
higher temperature --- reproducing the experimental behaviour instead of tracking
the density monotonically. The numerical consequence is that $\rhobar$ is no
longer temperature-independent: any scanner that caches the
$\rhobar(\delta,\Delta_\eff)$ table must rebuild it per temperature with this
atom (Sec.~\ref{sec:dopp}).

\subsection{Liouville space (vectorization)}
\label{sec:liouville}

Equation~\eqref{eq:gksl} is linear in $\rho$, so stacking the $n\times n$ matrix
$\rho$ into a length-$n^2$ vector turns the whole right-hand side into a single
matrix --- the Liouvillian $\Lcal$. The code uses row-major vectorization,
$\vecop(\rho)_{n i+j}=\rho_{ij}$, for which the commutator becomes
(\code{core.comm\_super}):
\begin{equation}
\vecop\big([H,\rho]\big)=\big(H\otimes \mathbb{1}-\mathbb{1}\otimes H^{T}\big)\vecop(\rho),
\qquad
\mathcal{C}[H]\equiv-\ii\big(H\otimes \mathbb{1}-\mathbb{1}\otimes H^{T}\big).
\label{eq:commsuper}
\end{equation}
The dissipator~\eqref{eq:lindblad} vectorizes likewise; each term
$L\rho L^\dagger$ becomes $L\otimes L^{*}$, and the anticommutators become
$\tfrac12(L^\dagger L\otimes\mathbb{1}+\mathbb{1}\otimes(L^\dagger L)^{T})$
(\code{atoms.\_build\_lindblad}). The full base Liouvillian is
\begin{equation}
\Lcal_0=\mathcal{C}[H_0]+\Dcal,
\label{eq:liouvillian}
\end{equation}
(\code{core.build\_liouvillian}). Finding the steady state is then the linear
algebra problem $\Lcal_0\,\vecop(\rho)=0$ subject to $\Tr\rho=1$; the trace
condition replaces one redundant row of $\Lcal_0$ (row $0$, the
$\rho_{g_1g_1}$ equation) by $\sum_i\rho_{ii}=1$. This ``replace a row with the
trace'' construction recurs in every solver in \code{core.py}.

% =====================================================================
\section{Thermal motion: the Doppler average}
\label{sec:dopp}
% =====================================================================

\subsection{Velocity classes}

At $T\sim120\,^\circ$C the atoms move fast enough that the Doppler shift $k v$
($\sigma_v=\sqrt{k_B T/m}\sim170\,$m/s gives $k\sigma_v/2\pi\sim200\,$MHz) dwarfs
the natural linewidth. The vapour is therefore a Maxwell ensemble of velocity
classes (\code{doppler.velocity\_grid}):
\begin{equation}
v\in[-v_{\max},v_{\max}],\quad
v_{\max}=\lceil c_\sigma\,\sigma_v\rceil,\quad
w(v)\propto \exp\!\Big(-\frac{v^2}{2\sigma_v^2}\Big),\quad \sum_v w(v)=1,
\end{equation}
with cutoff $c_\sigma=3$ standard deviations and a step $\dd v$ of a few m/s. Each
class sees both lasers Doppler-shifted by $kv$ (the pump and the near-degenerate
seed/conjugate share essentially the same $k$, since they differ only by the GHz
hyperfine scale $\ll$ the optical frequency).

\subsection{The \boldmath$\Delta_\eff$ trick}

The crucial observation (\code{doppler.py} header, README ``Speed'') is that the
\emph{only} place the one-photon detuning $\Delta$ and the velocity $v$ enter the
Hamiltonian is through the excited-state diagonal. Define the effective one-photon
detuning
\begin{equation}
\Delta_\eff=\Delta-k v.
\end{equation}
Then the velocity-dependent Liouvillian is a \emph{rigid shift} of one template,
\begin{equation}
\Lcal_0(\Delta_\eff)=\Lcal_0(0)-\Delta_\eff\,S_v,
\qquad
S_v=\mathcal{C}\Big[\textstyle\sum_{e\in\text{excited}}\ketbra{e}{e}\Big],
\label{eq:deffshift}
\end{equation}
where $S_v$ is the commutator super-operator of the projector onto the excited
manifold (\code{atoms.\_build\_velocity\_shift}). Consequently:
\begin{itemize}[leftmargin=1.6em]
  \item In the \emph{Hamiltonian}, temperature enters only through the Maxwell
  weights $w(v)$, and the detuning $\Delta$ only as an offset of the
  $\Delta_\eff$ axis.
  \item For a \emph{fixed dissipator} the susceptibility table
  $\rhobar(\delta,\Delta_\eff)$ is therefore a function of
  $(\delta,\Delta_\eff)$ alone, with no explicit $T$ or absolute-$\Delta$
  dependence.
  \item The collisional decoherence of Sec.~\ref{sec:collisional} puts $N(T)$
  into the dissipator, so the table is now $T$-dependent through
  $\gamma_{\text{opt}}(N)$ and $\gamma_{gg}(T,N)$: \emph{one table per
  temperature}, still reusable across all $\Delta$ and all velocity classes at
  that temperature.
\end{itemize}
The $\Delta_\eff$ sampling axis is built (\code{doppler.build\_Delta\_eff\_axis})
to span $\Delta-k v$ across a spectrum call's velocity grid, so the later average
is a cheap index interpolation rather than a re-solve. The dedicated
$(\Delta,T)$ squeezing map (\code{analysis/scan\_squeezing\_delta\_T.py}) builds,
per temperature, a master $\Delta_\eff$ axis from the exact union of all velocity
samples needed by every $\Delta$ in the scan and evaluates
$\rhobar(\delta,\Delta_\eff)$ once on that master grid with that temperature's
collisional atom; each map point then only gathers the relevant columns and
applies the Maxwell weights.

% =====================================================================
\section{The multimode steady state (truncated Floquet)}
\label{sec:floquet}
% =====================================================================

\subsection{Why the steady state is not static}

In ordinary absorption one weak field sits inside $H_0$ and the steady state is a
single static $\rho$. FWM is different: the seed and the conjugate are
\emph{different frequencies}, separated by twice the ground hyperfine splitting,
and they couple through the ground-state coherence $\rho_{g_1g_2}$, which therefore
\emph{oscillates} at the beat frequency
\begin{equation}
\Omega_{\text{beat}}=\Omega_{\HF}-\text{branch}\cdot\delta,
\qquad \Omega_{\HF}=2\pi\nu_{\HF}.
\label{eq:beat}
\end{equation}
The Hamiltonian is thus time periodic. We split it into the static part
$H_0$ (pump $+$ one weak field, Eq.~\eqref{eq:H0}) and a \emph{sideband} part
that oscillates at $\Omega_{\text{beat}}$ (\code{sideband\_hamiltonian}):
\begin{equation}
H(t)=H_0+H_p\,e^{-\ii\Omega_{\text{beat}}t}+H_p^\dagger\,e^{+\ii\Omega_{\text{beat}}t}.
\label{eq:Ht}
\end{equation}
$H_p$ contains the \emph{other} weak field plus the pump on the complementary leg;
its one-sided form ($\ketbra{e}{g}$ only, \code{\_add\_sideband\_drive}) is the
positive-frequency component of that drive.

\subsection{Three-mode ansatz and truncation}

Because the drive is weak, the response is dominated by the carrier and the first
sidebands. Expand
\begin{equation}
\rho(t)=\rho_0+\rho_{+1}\,e^{-\ii\Omega_{\text{beat}}t}+\rho_{-1}\,e^{+\ii\Omega_{\text{beat}}t}+\dots
\end{equation}
and keep $|n|\le1$ (a three-mode truncated Floquet expansion). Substituting into
the vectorized master equation and matching harmonics gives, with
$C_p\equiv\mathcal{C}[H_p]$ and $C_m\equiv\mathcal{C}[H_p^\dagger]$
(\code{sideband\_template}):
\begin{align}
(\Lcal_0+\ii\Omega_{\text{beat}})\,\rho_{+1}&=-C_p\,\rho_0, \label{eq:sbplus}\\
(\Lcal_0-\ii\Omega_{\text{beat}})\,\rho_{-1}&=-C_m\,\rho_0, \label{eq:sbminus}\\
\Lcal_0\,\rho_0&=-\big(C_p\,\rho_{-1}+C_m\,\rho_{+1}\big). \label{eq:carrier}
\end{align}
Equation~\eqref{eq:carrier} is the carrier steady state \emph{dressed} by the
sidebands feeding back into it.

\subsection{Feedback elimination}

Solving Eqs.~\eqref{eq:sbplus}--\eqref{eq:sbminus} for the sidebands and
substituting into~\eqref{eq:carrier} eliminates $\rho_{\pm1}$, leaving a single
effective equation for the carrier (\code{core.floquet\_solve}):
\begin{equation}
\boxed{\;
\Big[\Lcal_0
-\;C_p\,(\Lcal_0-\ii\Omega_{\text{beat}})^{-1}\,C_m
-\;C_m\,(\Lcal_0+\ii\Omega_{\text{beat}})^{-1}\,C_p\Big]\rho_0=0.
\;}
\label{eq:Aeff}
\end{equation}
The two correction terms are exactly the code's \code{minus\_feedback} and
\code{plus\_feedback}; the bracket is \code{A\_eff}. Trace-normalizing (row $0$
$\to\sum_i\rho_{0,ii}=1$) and solving gives $\rho_0$; the upper sideband then
follows directly,
\begin{equation}
\rho_{+1}=-(\Lcal_0+\ii\Omega_{\text{beat}})^{-1}\,C_p\,\rho_0,
\label{eq:rhop1}
\end{equation}
reusing the factor already computed for \eqref{eq:Aeff} rather than solving a
second system. Physically, $\rho_0$ holds the optical coherence at the seed
frequency (the probe response) and $\rho_{+1}$ holds the coherence at the
conjugate frequency (the generated conjugate) --- the two outputs we need.

\paragraph{Compiled grid solve.} Equation~\eqref{eq:deffshift}, together with the
affine $\delta$ dependence of the ground-state Hamiltonian, lets the FWM grid be
assembled from constant matrices:
\begin{equation}
\Lcal_0(\delta,\Delta_\eff)
=\Lcal_0(0,0)+\delta\,C_\delta-\Delta_\eff\,S_v,
\qquad
\Omega_{\text{beat}}(\delta)=\Omega_{\HF}-\text{branch}\cdot\delta .
\end{equation}
When Numba is available, \code{kernels.floquet\_chi\_grid} evaluates the full
$(\delta,\Delta_\eff)$ grid in a fused parallel loop: assemble the local
Liouvillian, factor the two sideband matrices, form $A_{\text{eff}}$, solve the
trace-normalized carrier, recover $\rho_{+1}$, and contract directly with the
CG-weighted readout vectors. The NumPy path follows the same equations with a
per-$\delta$ loop over \code{core.floquet\_solve}; \code{tests/test\_kernels.py}
pins the two paths to the same numerical result.

% =====================================================================
\section{The susceptibility matrix}
\label{sec:chi}
% =====================================================================

\subsection{What \boldmath$\rhobar$ means}

We want the linear response of the two optical polarizations (one per Raman leg)
to the two weak fields. Define the dimensionless susceptibility matrix by
\begin{equation}
\begin{pmatrix}P_s\\[2pt]P_c\end{pmatrix}
=
\begin{pmatrix}\rhobar_{ss}&\rhobar_{sc}\\[2pt]\rhobar_{cs}&\rhobar_{cc}\end{pmatrix}
\begin{pmatrix}\Omega_s\\[2pt]\Omega_c\end{pmatrix},
\label{eq:chimatrix}
\end{equation}
where $P_s$ ($P_c$) is the CG-weighted optical coherence read out on the
seed (conjugate) leg, and $\Omega_s$ ($\Omega_c$) is the seed (conjugate) Rabi
frequency. The polarization readout is (\code{\_polarization\_coherence}):
\begin{equation}
P_{\text{leg }g}=\sum_{e\in\text{excited}}\sqrt{3 C_F^2(F_g,F_e)}\;\rho_{e g},
\label{eq:Pleg}
\end{equation}
the same CG weighting as the drive, so absorption and emission are described by one
consistent dipole. Because the response is linear in the \emph{weak} fields (the
pump is the strong field, already inside $H_0$ and the sidebands), $\rhobar$ is
\emph{independent} of the magnitude of the reference field that probes it --- the
code simply divides the measured coherence by the reference Rabi
$\Omega_{\text{ref}}$.

\subsection{Two solves per detuning}

Each entry of the $2\times2$ matrix is a response of one leg to one field, so two
independent Floquet solves at each two-photon detuning $\delta$ extract all four
(\code{chi\_matrix\_table}):
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{Solve 1 --- seed drive only} ($\Omega_c=0$ in the sidebands,
  $\Omega_s=\Omega_{\text{ref}}$ in $H_0$). The carrier coherence on the seed leg
  gives $\rhobar_{ss}=P_s/\Omega_{\text{ref}}$; the upper-sideband coherence on the
  conjugate leg gives $\rhobar_{cs}=P_c/\Omega_{\text{ref}}$ --- this is the seed
  $\to$ conjugate \emph{generation}.
  \item \textbf{Solve 2 --- conjugate seed only} ($\Omega_c=\Omega_{\text{ref}}$ in
  the sidebands, $\Omega_s=0$). Symmetrically yields $\rhobar_{sc}$ and
  $\rhobar_{cc}$.
\end{itemize}
The off-diagonal $\rhobar_{cs},\rhobar_{sc}$ are the parametric coupling that does
not exist without the pump; they are the heart of the four-wave mixing.
\code{chi\_matrix\_table} accepts the atom model as an argument
(\code{atom=}): \code{compute\_spectrum} passes the per-temperature collisional
atom of Sec.~\ref{sec:collisional}, while the default natural-linewidth
\code{ATOM} keeps the kernel/NumPy regression callers unchanged. In the
compiled path, the two physical solves correspond to two calls to
\code{kernels.floquet\_chi\_grid}: one with the seed reference field and one with
the conjugate reference field. Each call fills the whole
$(\delta,\Delta_\eff)$ grid and returns the probe-leg and conjugate-leg
coherences already contracted with the proper CG readout weights.

\subsection{Detuning conventions and the two Raman branches}
\label{sec:branch}

The probe is scanned in laboratory frequency, referenced to the
$F=2\!\to\!F'=3$ line. The two Raman resonances sit at
(\code{branch\_center\_GHz})
\begin{equation}
f^{(\pm)}_{\text{Raman}}=\Delta_{\text{GHz}}\pm\nu_{\HF},
\end{equation}
and the two-photon detuning of a probe point is
(\code{two\_photon\_detuning\_from\_probe\_scan})
\begin{equation}
\delta=2\pi\big(f_{\text{probe}}-f^{(\text{branch})}_{\text{Raman}}\big).
\end{equation}
The beat frequency~\eqref{eq:beat} that drives the Floquet expansion uses this same
$\delta$. \textbf{The two branches are separate physical channels} (separate
probe/conjugate mode pairs): the standard squeezing operating point is the
red-detuned seed on the $(-)$ branch. Adding the two branches' $\rhobar$ matrices
into one propagation is wrong --- it manufactures spurious high-gain features (the
README documents an artefact near $+70\,$MHz TPD). The wide ``both branches'' view
computes branch $-1$ and $+1$ independently and overlays them
(\code{full\_spectrum}).

% =====================================================================
\section{Doppler averaging of the susceptibility}
% =====================================================================

The single-velocity table $\rhobar(\delta,\Delta_\eff)$ is averaged over the
Maxwell distribution to give the macroscopic susceptibility seen by the beam.
Using $\Delta_\eff=\Delta-kv$ and the uniform $\Delta_\eff$ axis, the average is a
weighted linear interpolation (\code{doppler.doppler\_average}):
\begin{equation}
\overline{\rhobar}(\delta;\Delta,T)
=\sum_v w(v)\,\rhobar\big(\delta,\;\Delta-kv\big).
\label{eq:doppleravg}
\end{equation}
This is a \emph{coherent average of the complex susceptibility}: amplitude and
phase from each velocity class add before the field propagates, which is the
correct order of operations for a dispersive parametric medium (the alternative ---
averaging intensities --- would discard the inhomogeneous phase that sets the
gain). The interpolation indices and fractions are precomputed once
(\code{interpolation\_weights}) and applied to every row of the table
(\code{apply\_doppler\_average}), so the average costs one matrix multiply. In the
$(\Delta,T)$ squeezing-map script, the per-temperature master $\Delta_\eff$ axis
is the exact union of all requested $\Delta-kv$ samples at that temperature. The
Doppler step therefore becomes a direct column gather followed by the same
weighted sum, with no interpolation error and no additional Floquet solve.

% =====================================================================
\section{Maxwell--Bloch propagation: susceptibility \boldmath$\to$ gain}
\label{sec:mb}
% =====================================================================

\subsection{Coupled amplitude equations and phase conjugation}

In a dilute medium the slowly varying seed and conjugate Rabi amplitudes obey
coupled propagation equations along the cell ($z$ direction). The combination that
closes is $(\Omega_s,\Omega_c^{*})^{T}$ --- note the \emph{complex conjugate} of
the conjugate field, the signature of a phase-conjugating ($\chi^{(3)}$) process:
\begin{equation}
\frac{\dd}{\dd z}\begin{pmatrix}\Omega_s\\[2pt]\Omega_c^{*}\end{pmatrix}
= M\begin{pmatrix}\Omega_s\\[2pt]\Omega_c^{*}\end{pmatrix},
\qquad
M=\frac{\ii}{2}
\begin{pmatrix}
 k_s\,\chi_{ss} & k_s\,\chi_{sc}\\[4pt]
 -k_c\,\chi_{cs}^{*} & -k_c\,\chi_{cc}^{*}
\end{pmatrix}.
\label{eq:Mmatrix}
\end{equation}
The diagonal terms are self-action (gain/absorption and dispersion on each beam);
the off-diagonal terms are the parametric coupling that converts seed into
conjugate and back. The conjugation and the minus signs in the second row are what
make $M$ generate \emph{amplification} (a non-unitary, energy-non-conserving
transfer) rather than mere rotation.

\subsection{Seeded longitudinal phase mismatch}

The seeded propagation matrix can carry a rotating-frame longitudinal mismatch
term:
\begin{equation}
M_{\Delta k}=
M+
\frac{\ii\,\Delta k_z}{2}
\begin{pmatrix}
1 & 0\\
0 & -1
\end{pmatrix}.
\label{eq:Mdeltak}
\end{equation}
For the seeded double-$\Lambda$ geometry the vacuum mismatch used in Balanced
resolution is
\begin{equation}
\Delta k_z^{(0)}(\delta)
=2k_p-\big(k_s(\delta)+k_c(\delta)\big)\cos\theta,
\qquad
\theta=0.32^\circ,
\label{eq:seededDk}
\end{equation}
where the reference angle is the Sim \emph{et al.}\ pump--probe crossing angle,
$k_s$ is set by the scanned seed frequency, and energy conservation sets
$k_c=2k_p-k_s$ in angular-frequency units. This is cheap: it is one vector over
the already-computed probe axis, and no OBE/Floquet solve is repeated.

High fidelity adds two post-OBE refinements while keeping the expensive
susceptibility table fixed:
\begin{enumerate}[leftmargin=2.2em]
  \item A conservative refractive correction,
  \[
  k_s\to n_s k_s,\qquad k_c\to n_c k_c,\qquad
  n_{s,c}\simeq1+\frac{1}{2}\Real\chi_{ss,cc},
  \]
  with the correction clipped to the dilute-vapour range so resonant dispersive
  excursions do not create artificial phase-matching runaways.
  \item A segmented propagation profile. The cell is sliced into 16 segments; the
  diagonal self-action remains fixed, while the off-diagonal parametric coupling
  is smoothly attenuated by a scalar profile inferred from the median absorptive
  optical depth. This is a propagation-only approximation, not a new atomic solve.
\end{enumerate}

The \code{phase\_detail} option selects how much of this propagation physics is
included. \code{PHASE\_LEGACY} sets $\Delta k_z=0$ for regression-preserving
direct calculations; the app's Fast and Balanced presets use
Eq.~\eqref{eq:seededDk}; High fidelity uses the refractive and segmented
refinements above; Ultra uses the dynamic-depletion propagation described next.

\subsection{Ultra propagation: dynamic pump budget, overlap, diagnostics}
\label{sec:ultraprop}

Ultra (\code{\_ultra\_segmented\_gain}) slices the cell into 64 segments and
propagates the two-component amplitude explicitly, segment by segment:
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{Refractive phase matching, one pass.} The refractive indices
  $n_{s,c}(\chi)$ do not depend on $\Delta k_z$, so there is no fixed point to
  iterate: a single refractive update of Eq.~\eqref{eq:seededDk} is exact
  (\code{ULTRA\_PHASE\_ITERATIONS}\,$=1$). The reported diagnostic is the
  vacuum$\to$refractive shift of $\Delta k_z$, not the last delta of a no-op
  loop.
  \item \textbf{Gaussian beam overlap.} The pump--probe crossing angle
  separates the beams along the cell; the off-diagonal coupling in each
  segment is weighted by the crossed-Gaussian amplitude overlap
  $\exp\!\big[-z^2\tan^2\theta/(w_p^2+w_s^2)\big]$, normalized to its maximum
  (\code{\_gaussian\_overlap\_profile}).
  \item \textbf{Segment-wise pump budget.} After each segment the added seed
  power and the generated conjugate power are subtracted from the remaining
  pump budget, and the parametric (off-diagonal) coupling of the next segment
  is scaled by $\sqrt{P_{\text{remaining}}/P_{\text{pump}}}$ --- an approximate
  \emph{dynamic} pump depletion along $z$, on top of the global Manley--Rowe
  cap. The minimum remaining pump along the scan is reported.
  \item \textbf{In-cell loss is not applied here.} The absorptive segment
  profile attenuates only the parametric coupling; the in-cell \emph{loss} is
  charged once, downstream, as the beamsplitter vacuum admixture of
  Sec.~\ref{sec:hardened}. (Applying it here as field attenuation too
  double-counted the same optical depth in an earlier Ultra version.)
  \item \textbf{Zeeman diagnostic.} The 24-level $^{85}$Rb $D_1$ Zeeman
  manifold is built and its CG sums are checked against the lumped $3C_F^2$
  strengths (consistency $\equiv1.0$ by construction). It is reported as a
  diagnostic only --- deliberately \emph{not} multiplied into the coupling ---
  and the full 24-level Floquet scan is not run in Ultra v1.
\end{itemize}

\subsection{From dimensionless to physical susceptibility}
\label{sec:coupling}

The dimensionless $\overline{\rhobar}$ becomes a physical susceptibility through
the medium coupling constant (\code{observables.gain\_from\_chi}):
\begin{equation}
\chi_{xy}= -\,\frac{2 N\,|d_\eff|^2}{\varepsilon_0\hbar}\,\overline{\rhobar}_{xy},
\qquad
|d_\eff|^2=\ell_s\,\varrho_{\text{norm}}\,|d_{D_1}|^2,
\label{eq:chiphys}
\end{equation}
where $N$ is the number density and $d_{D_1}$ the $D_1$ reduced dipole. The sign
makes a passive transition absorptive ($\Imag\chi>0$ on resonance). The effective
dipole factorizes into a \emph{first-principles} normalization and a
\emph{residual} calibration:
\begin{itemize}[leftmargin=1.6em]
  \item $\varrho_{\text{norm}}=p_F/[2(2I+1)]$
  (\code{fwm.physical\_coupling\_norm}) is the thermal ground-population
  fraction of the probed manifold, $p_F=(2F+1)/12$, divided by the ground
  Zeeman degeneracy $2(2I+1)=12$ ($I=5/2$). The AutoOD-validated absorption path
  ($<0.1\,\%$ against measured cell transmission) carries exactly this factor;
  the lumped 4-level solve instead normalizes $\Tr\rho=1$ over just four levels
  and multiplies by the \emph{total} atomic density, so the factor would
  otherwise be missing from the macroscopic scale. The seeded probe couples from
  $F=3$ on the standard $(-)$ Raman branch, giving
  $\varrho_{\text{norm}}=\tfrac{7}{12}\cdot\tfrac{1}{12}
  =\tfrac{7}{144}\approx0.0486$
  (and $\tfrac{5}{144}\approx0.0347$ on the $(+)$ branch). Quantitatively, this
  is the factor the old hand-tuned fudge ``$\text{line\_strength}=0.05$'' was
  standing in for.
  \item $\ell_s$ (\code{LINE\_STRENGTH\_FACTOR}; the app's Advanced
  ``line-strength factor'') is the dimensionless \emph{residual} left on top,
  engine default $1.0$. The seeded app default is $\ell_s=0.74$, anchored so
  that Ultra fidelity reproduces the measured $G_s\approx14$ and $-7.8\,$dB at
  the Sim \emph{et al.}\ operating point; the departure from $1$ stands for the
  $m$-sublevel participation / beam-geometry reduction not yet derived from
  first principles. It is not an experimentally tunable quantity, which is why
  the app parks it under \emph{Advanced}.
\end{itemize}
The bookkeeping is now clean: the genuine hyperfine CG structure is inside
$\rhobar$ (drive $\times$ readout $=3C_F^2$ per leg), the population/degeneracy
normalization is $\varrho_{\text{norm}}$, and the density is the validated CRC
value --- the same normalization chain that reproduces measured linear
absorption. $\ell_s$ rescales the \emph{effective} $|d|^2$ in propagation only;
it does \emph{not} touch the OBE solve (where $\Gamma$ and $\Omega$ are fixed by
$I_{\text{sat}}$).

\subsection{The transfer matrix}

Because $M$ is (per detuning point) a constant $2\times2$ matrix over the cell
length $L$, the propagation is solved exactly by the matrix exponential
\begin{equation}
T=\exp(M L),\qquad
\begin{pmatrix}\Omega_s\\\Omega_c^{*}\end{pmatrix}_{\!\!z=L}
=T\begin{pmatrix}\Omega_s\\\Omega_c^{*}\end{pmatrix}_{\!\!z=0}.
\end{equation}
The $2\times2$ exponential is evaluated in closed form
(\code{core.matrix\_exp\_2x2}; derivation in Appendix~\ref{app:expm}) for the whole
batch of detuning points at once. The power gains are
\begin{equation}
G_s=|T_{00}|^2 \quad(\text{seed transmission}),
\qquad
G_c=|T_{10}|^2 \quad(\text{conjugate generated from a seed at port 0}).
\label{eq:gains}
\end{equation}
With a seed in and vacuum in the conjugate port, $G_s$ is the amplified seed power
and $G_c$ is the newly generated conjugate power. In the high-gain parametric
regime the twin-beam relation $G_c\simeq G_s-1$ holds (each created conjugate
photon accompanies one added seed photon).

\paragraph{The exponential-density trap.} Since $M\propto N$ (Eq.~\eqref{eq:chiphys})
and $T=e^{ML}$, the gain grows \emph{exponentially} with density, i.e.\
super-exponentially with temperature through $N(T)$. This is physically correct
only while the pump is undepleted; pushed too far it over-amplifies, which the next
section caps.

\paragraph{Overflow clamp.} Numerically, \code{core.matrix\_exp\_2x2} clips the
\emph{real part} of the exponent arguments $cL$ and $sL$ at $\pm350$ --- two such
factors multiply, and float64 $e^{x}$ overflows near $x\simeq709$. Without the
clamp, a runaway linear gain returns \code{inf}, and the downstream
pump-depletion saturation then evaluates $\infty/(1+\infty)=\text{NaN}$,
silently poisoning the whole gain/squeezing curve. With it, an unphysical linear
gain stays huge-but-finite and is capped to the energy bound by the saturation.
The clamp is a strict no-op in the validated regime ($|\Real(cL)|\ll350$).

% =====================================================================
\section{Pump depletion: the Manley--Rowe cap}
% =====================================================================

The propagation above is linear in the \emph{undepleted} pump, so at high density
it can predict $G_s P_{\text{seed}}\gg P_{\text{pump}}$ --- more output power than
the pump can supply. Non-degenerate FWM is a Manley--Rowe process: two pump photons
make one signal photon and one conjugate photon, so at full conversion the added
signal power and the generated conjugate power each approach half the pump power.
The code enforces this energy bound with a smooth homogeneous-saturation form
(\code{observables.pump\_depletion\_saturation}):
\begin{equation}
G_s^{\text{sat}}=1+\frac{G_s-1}{1+(G_s-1)\,P_{\text{seed}}/P_{\text{cap}}},
\qquad
G_c^{\text{sat}}=\frac{G_c}{1+G_c\,P_{\text{seed}}/P_{\text{cap}}},
\qquad
P_{\text{cap}}=\frac{P_{\text{pump}}}{2}.
\label{eq:manleyrowe}
\end{equation}
Three properties make this the right minimal fix:
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{Small-signal limit.} When $(G-1)P_{\text{seed}}\ll P_{\text{cap}}$
  the denominators are $\approx1$ and the validated gain is untouched.
  \item \textbf{Energy ceiling.} As $G\to\infty$, $(G_s^{\text{sat}}-1)P_{\text{seed}}\to
  P_{\text{cap}}$ and $G_c^{\text{sat}}P_{\text{seed}}\to P_{\text{cap}}$ --- exactly the
  half-pump bound.
  \item \textbf{Twin-beam relation preserved.} $(G_s-1)$ and $G_c$ saturate
  identically, so $G_c^{\text{sat}}\simeq G_s^{\text{sat}}-1$ still holds and the
  squeezing formula stays consistent.
\end{itemize}
The pipeline keeps the pre-saturation small-signal peak for diagnostics and flags
the operating point as ``pump-depletion limited'' when the small-signal gain
exceeds the cap by more than 10\,\% (\code{compute\_spectrum},
\code{\_seeded\_observables}).

% =====================================================================
\section{Quantum observables of the twin beam}
% =====================================================================

\subsection{Intensity-difference squeezing}
\label{sec:ids}

The twin beams are correlated, so the noise in $\hat n_s-\hat n_c$ can fall below
the shot-noise level. For the lossless seeded parametric amplifier the correct
starting point is the two-mode Bogoliubov transformation
$\hat a_s\to u\,\hat a_s+v\,\hat a_c^\dagger$,
$\hat a_c\to u\,\hat a_c+v\,\hat a_s^\dagger$ with $|u|^2-|v|^2=1$, i.e.\
$G_s=|u|^2$, $G_c=|v|^2=G_s-1$. This transformation \emph{conserves the
photon-number difference}: $\hat n_s-\hat n_c$ commutes with the two-mode squeeze
operator, so
\begin{equation}
\big(\hat n_s-\hat n_c\big)_{\text{out}}=\big(\hat n_s-\hat n_c\big)_{\text{in}}.
\label{eq:conserved}
\end{equation}
With a bright coherent seed in port $s$ (Poissonian,
$\mathrm{Var}=\langle n\rangle_{\text{in}}$) and vacuum in port $c$, the output
difference noise is just the \emph{input} seed shot noise, while the shot-noise
reference of the output beams is
$\langle\hat n_s+\hat n_c\rangle_{\text{out}}=(G_s+G_c)\langle n\rangle_{\text{in}}$.
Generalizing the prefactor to unequal gains (which the propagation over
absorptive/dispersive $\chi$ produces), the ideal noise implemented in
\code{observables.ideal\_twin\_beam\_noise} is
\begin{equation}
\boxed{\;
S_{\text{ideal}}=\frac{(G_s-G_c)^2}{G_s+G_c}
\;}
\qquad\xrightarrow[\ G_s-G_c\,=\,1\ ]{\text{lossless}}\qquad
\frac{1}{G_s+G_c}=\frac{1}{2G_s-1},
\label{eq:Sideal}
\end{equation}
the standard result of the FWM squeezing literature (Sim \emph{et al.};
McCormick \emph{et al.}, \emph{Opt.\ Lett.}\ \textbf{32}, 178 (2007)).
At unity gain $S_{\text{ideal}}=1$ (shot noise); at high gain
$S_{\text{ideal}}\to1/(2G_s-1)\to0$. The value is clipped to $[0,1]$ so that
passive-loss regions of a probe scan --- no parametric gain, $G_s-G_c$ arbitrary
--- cannot spuriously report sub-shot-noise values.

\begin{remark}[The corrected formula]
Until 2026-07-01 the code used
$S_{\text{old}}=\big(G_s+G_c-2\sqrt{G_sG_c-1}\big)/(G_s+G_c)$. The two
expressions agree only at $(G_s,G_c)=(2,1)$. On the twin-beam line $G_c=G_s-1$
the old numerator is a catastrophic cancellation whose expansion gives
$S_{\text{old}}\simeq\tfrac{5}{2}/(2G_s-1)^2$ instead of $1/(2G_s-1)$ --- a
spurious extra squeezing factor $\approx(2G_s-1)/2.5$ that grows with gain
(at $G_s\!\sim\!15$, about $+10\,$dB of fictitious squeezing). Under normal
operation the error was masked because the detection floor $(1-\eta)$ dominates
$S$, but it corrupted every ideal-detection study: re-evaluating the same
$(\Delta,T)$ ideal-scan data gives $-84.7$\,dB with the old formula versus
$-31.3$\,dB with the corrected one (squeezing report v4). The pinned regression
baseline was re-captured with the corrected formula.
\end{remark}

Symmetric detection efficiency $\eta$ on both arms mixes in uncorrelated vacuum,
\begin{equation}
S(\eta)=\eta\,S_{\text{ideal}}+(1-\eta),
\qquad
S_{\text{dB}}=10\log_{10}S,
\label{eq:Seta}
\end{equation}
with $\eta=\text{QE}\,(1-\text{loss})$. Equation~\eqref{eq:Seta} immediately gives
the squeezing \emph{floor}: since $S_{\text{ideal}}\!\ge\!0$,
\begin{equation}
S_{\text{dB}}\ \ge\ 10\log_{10}(1-\eta),
\label{eq:floor}
\end{equation}
a hard limit set purely by loss. At the reference detection chain
($\text{QE}=0.92$, $5.5\,\%$ optical loss, $\eta=0.8694$) the floor is
$-8.84\,$dB. (The 2-D $\xi(\Delta,T)$ scan confirms the model saturates exactly
at this plateau once the gain is large enough --- gain saturation and the loss
floor together cap the achievable squeezing.) Negative $S_{\text{dB}}$ means
squeezed below shot noise; the regression-pinned legacy scan at the Sim
\emph{et al.}\ reference setting reaches $S_{\text{dB}}\approx-8.69\,$dB at
$G_s\approx32$ (Sec.~\ref{sec:pipeline}).

\subsection{Beyond symmetric loss: the balanced two-arm noise model}
\label{sec:balanced}

Real detection is rarely symmetric: the two arms can have different total
transmissions, and the reference photocurrent can be electronically weighted.
The generalized noise (\code{observables.balanced\_twin\_beam\_noise}) is the
variance of the weighted difference photocurrent, normalized to the
coherent-state shot noise of the same combination:
\begin{equation}
D=I_s-w\,I_c,
\qquad
S=\frac{\mathrm{Var}\,D}{\langle I_s\rangle+w^2\langle I_c\rangle},
\qquad
\langle I_s\rangle=\eta_s G_s,\quad \langle I_c\rangle=\eta_c G_c,
\label{eq:balancedS}
\end{equation}
with total arm efficiencies $\eta_s,\eta_c$ (cell transmission $\times$ path
loss $\times$ QE) and
\begin{equation}
\mathrm{Var}\,D=\eta_sG_s+w^2\eta_cG_c-2w\,\eta_s\eta_c\,\mathrm{cov}_0,
\qquad
\mathrm{cov}_0=\tfrac12\Big[(G_s+G_c)-(G_s-G_c)^2\Big],
\label{eq:cov0}
\end{equation}
where $\mathrm{cov}_0$ is the source intensity covariance consistent with the
conserved-difference result: at $w=1$, $\eta_s=\eta_c=\eta$ the model reduces
\emph{exactly} to Eqs.~\eqref{eq:Sideal}--\eqref{eq:Seta}, and on the lossless
twin-beam line $\mathrm{cov}_0=G_c$. It is clipped to the Cauchy--Schwarz bound
$\sqrt{G_sG_c}$ (and to $\ge0$). The reference weight $w$ may be a scalar
electronic gain or one of three lab conventions: \code{raw} ($w=1$), \code{dc}
($w=\langle I_s\rangle/\langle I_c\rangle$, cancelling the mean photocurrents
--- what a balanced detector with a DC servo does), and \code{shot}
($w=\sqrt{\langle I_s\rangle/\langle I_c\rangle}$, minimizing the normalized
quantum noise at fixed detected powers). An optional \code{source\_noise} input
replaces the ideal source covariance with a measured/inferred value (used by
paper-anchored resonant-FWM readouts elsewhere in GABES), and additive
technical excess-noise terms enter as $S\to S+\mathcal{N}_{\text{tech}}$. This
model is the noise engine of the hardened Ultra path
(Sec.~\ref{sec:hardened}).

\subsection{Twin-beam photon statistics}

In the same ideal-parametric spirit, the equal-time photon statistics follow from
the mean pair number $n=G_s-1$ (\code{observables.coincidence\_stats}):
\begin{equation}
g^{(2)}_{ss}=g^{(2)}_{cc}=2\ (\text{each arm thermal}),
\qquad
g^{(2)}_{sc}(0)=2+\frac{1}{n},
\qquad
R=\frac{[g^{(2)}_{sc}]^2}{g^{(2)}_{ss}\,g^{(2)}_{cc}}=\frac{(2+1/n)^2}{4}.
\label{eq:g2}
\end{equation}
The Cauchy--Schwarz ratio $R>1$ everywhere there is net gain is the nonclassical
photon-pair signature; $g^{(2)}_{sc}\to2$ at high flux and diverges at low flux.
These are valid only where $G_s>1$ (set to NaN otherwise) and, like the squeezing
above, do not include quantum-Langevin propagation loss.

% =====================================================================
\section{The Ultra noise model: in-cell loss, asymmetric arms, pump scatter}
\label{sec:hardened}
% =====================================================================

The legacy/Balanced/Fine paths score squeezing with the symmetric formula
Eq.~\eqref{eq:Seta}. The Ultra path models the noise budget more honestly, in
two stages: a beamsplitter treatment of in-cell loss (pre-hardening Ultra), and
--- intrinsic since the v5 hardening --- an excess-noise model that charges the
penalties a gain-gap-only objective is blind to.

\subsection{In-cell loss as a beamsplitter}

Propagation loss inside the cell happens \emph{before} the detectors, so Ultra
applies it to the source noise as a vacuum admixture with transmission
$\tau_{\text{cell}}=e^{-\mathrm{OD}}$ (from the median absorptive optical depth
of the segment profile), then the detection efficiency, then any additive
technical noise (\code{observables.segmented\_loss\_noise\_squeezing\_dB}):
\begin{equation}
S_{\text{cell}}=\tau_{\text{cell}}\,S_{\text{ideal}}+(1-\tau_{\text{cell}}),
\qquad
S=\eta\,S_{\text{cell}}+(1-\eta)+\mathcal{N}_{\text{tech}}.
\label{eq:segloss}
\end{equation}
Deliberately, this loss is \emph{not} also applied as field attenuation inside
the segmented propagation --- doing both double-counts the same optical depth
(an early Ultra bug). The technical-noise channels
($\mathcal{N}_{\text{tech}}$: seed excess noise, pump scatter, EOM residual)
were zero-default placeholders in this stage; Eq.~\eqref{eq:segloss} with
$\mathcal{N}_{\text{tech}}=0$ is what \code{excess\_noise\_model=False} still
computes, kept for A/B against pre-hardening results.

\subsection{What a gap-only objective is blind to}

$S_{\text{ideal}}=(G_s-G_c)^2/(G_s+G_c)$ sees only the \emph{gain gap}: any
operating point with $G_s-G_c\approx1$ looks near-perfect regardless of how
large the gains are, where the two sidebands sit relative to the absorption
manifolds, and how much pump light reaches the detectors. Squeezing report v5
documents the consequence: with probe-only in-cell loss the $(\Delta,T)$
optimizer rewarded a near-resonance ultra-high-gain region ($G_s\sim636$ at an
in-cell OD reported as $\approx0$ --- an internal contradiction of the lumped
model) and pinned a spurious global optimum at $\Delta=-2.20\,$GHz,
$T=135\,^\circ$C.

\subsection{The hardened excess-noise model}

The hardened model (intrinsic to Ultra; \code{compute\_spectrum} argument
\code{excess\_noise\_model}, default on) adds exactly the two missing penalties.

\paragraph{(i) Arm-resolved linear absorption.} The lumped double-$\Lambda$
parametric susceptibilities $\chi_{ss},\chi_{cc}$ carry the FWM response but
\emph{not} the ordinary linear absorption a real beam suffers when its optical
frequency crosses the populated $F=2\!\to\!F'$ manifold --- the physics that
kills squeezing when a sideband lands near resonance (the conjugate in the
deep-red-$\Delta$ region, the probe in the deep-blue-$\Delta$ region; Sim
\emph{et al.}\ document the conjugate case). \code{fwm.\_arm\_linear\_od}
evaluates each arm's own linear optical depth from the AutoOD-validated
hyperfine absorption ($<0.1\,\%$ vs lab), summing the manifold components with
$F_g=2$ (branch $-1$) at the arm's own frequency; the conjugate frequency
follows from energy conservation, $f_c=2\Delta_{\text{GHz}}-f_{\text{probe}}$.
These enter as \emph{asymmetric arm transmissions} in the balanced model of
Sec.~\ref{sec:balanced}:
\begin{equation}
\eta_s^{\text{arm}}=\eta\,e^{-(\mathrm{OD}_{\text{seg}}
+\mathrm{OD}^{\text{lin}}_{s})},
\qquad
\eta_c^{\text{arm}}=\eta\,e^{-\mathrm{OD}^{\text{lin}}_{c}},
\label{eq:armloss}
\end{equation}
where $\mathrm{OD}_{\text{seg}}$ is the $\chi_{ss}$-based in-cell depth already
used by the segment profile and $\mathrm{OD}^{\text{lin}}_{s,c}$ are the
manifold linear depths (clipped to $[0,5]$). Asymmetric arm loss is precisely
what breaks the intensity-difference correlation, so parking either sideband on
a populated manifold is now charged its physical price.

\paragraph{(ii) Pump-scatter excess noise.} The pump's own linear optical depth
$\mathrm{OD}_p$ at its detuning $\Delta$ (same validated absorption path) fixes
how much pump power the vapour scatters; a residual fraction $\kappa$ of that
reaches the difference photocurrent (finite PBS extinction and spatial-filter
leakage):
\begin{equation}
\mathcal{N}_{\text{ps}}=\kappa\,\big(1-e^{-\mathrm{OD}_p}\big),
\qquad
\kappa=0.1\ \text{(default, the one residual coefficient)}.
\label{eq:pumpscatter}
\end{equation}
This is small far from resonance (the $+0.9\,$GHz window of Sim \emph{et al.})
and large when the pump sits near a populous manifold (the
$-2.1\,$GHz/$F{=}3$ region).

\paragraph{Assembled noise.} The Ultra squeezing is then the balanced two-arm
noise with the DC-cancelling weight and the pump scatter as additive excess
noise:
\begin{equation}
S_{\text{Ultra}}
=\mathrm{Var}\big(I_s-wI_c\big)\big/\big(\langle I_s\rangle+w^2\langle
I_c\rangle\big)\Big|_{\eta_s^{\text{arm}},\,\eta_c^{\text{arm}},\,w=\text{dc}}
\;+\;\mathcal{N}_{\text{ps}},
\qquad
S_{\text{dB}}=10\log_{10}S_{\text{Ultra}}.
\label{eq:hardenedS}
\end{equation}
The absorption channels carry \emph{no} tuning coefficient (they are validated
physics); $\kappa$ is the single residual.

\subsection{What hardening changes (squeezing report v5)}

\begin{itemize}[leftmargin=1.6em]
  \item \textbf{The extreme-gain artefact is removed.} The $(\Delta,T)$
  global-optimum gain drops $G_s=636\to27$; the v4 artefact coordinate
  ($-2.20\,$GHz, $135\,^\circ$C) is suppressed from $-8.83$ to $-5.41\,$dB by
  the conjugate arm's absorption.
  \item \textbf{The honest optimum stays slightly red-detuned.} The hardened
  global optimum is $\Delta=-2.0\,$GHz, $T=120\,^\circ$C, $\xi=-8.07\,$dB at
  $G_s=27$, with the Sim-side lobe ($\Delta=+1.3\,$GHz, $-7.79\,$dB) only
  $0.28\,$dB shallower --- near-degenerate. The residual red preference is real
  physics \emph{of this model}: the more populated $F=3$ ground manifold
  ($7/12$ vs $5/12$) and the spectrally cleaner far-red probe position. What the
  lumped steady-state model cannot contain is the optical-pumping depletion of
  $F=3$ when the pump parks near the $F=3$ manifold, which is the leading
  candidate for why the experiment prefers $+0.9\,$GHz.
  \item \textbf{Even the ideal-detection limit is capped.} With $\eta\to1$ the
  gap-only objective runs away to $-36.6\,$dB at the $G_s=636$ coordinate; the
  hardened model caps the ideal optimum at $-15.5\,$dB / $G_s=27$.
\end{itemize}
The legacy/Balanced/Fine squeezing formulas and the pinned regression baseline
are untouched by the hardening; \code{excess\_noise\_model=False} reproduces
pre-hardening Ultra, and a dict (e.g.\
\code{\{"pump\_scatter\_kappa": 0.2\}}) tunes $\kappa$.

% =====================================================================
\section{The seeded pipeline, end to end}
\label{sec:pipeline}
% =====================================================================

Collecting the pieces, \code{compute\_spectrum} (one one-photon detuning
$\Delta=2\pi\,D_{\text{GHz}}\!\cdot\!10^9$) runs:
\begin{enumerate}[leftmargin=2.2em]
  \item Convert powers to Rabi frequencies, Eq.~\eqref{eq:rabi}; CRC density
  $N(T)$, Eq.~\eqref{eq:crc}; coupling scale
  $\ell_s\varrho_{\text{norm}}$ (Sec.~\ref{sec:coupling}).
  \item Build the per-temperature collisional atom
  (\code{collisional\_atom}: $\gamma_{gg}(T,N)$, $\gamma_{\text{opt}}(N)$,
  Sec.~\ref{sec:collisional}).
  \item Build the probe-frequency scan axis around the chosen Raman branch, and map
  each point to a two-photon detuning $\delta$ (Sec.~\ref{sec:branch}).
  \item Build the velocity grid and the $\Delta_\eff$ axis (Sec.~\ref{sec:dopp}).
  \item Build the $\rhobar$ table on $(\delta,\Delta_\eff)$ with that atom
  (Secs.~\ref{sec:floquet}--\ref{sec:chi}); with Numba available this is the
  fused \code{kernels.floquet\_chi\_grid} path, otherwise the NumPy
  \code{core.floquet\_solve} reference path.
  \item Doppler-average each of the four $\rhobar$ entries,
  Eq.~\eqref{eq:doppleravg}.
  \item Maxwell--Bloch propagate $\to$ $G_s,G_c$,
  Eqs.~\eqref{eq:Mmatrix}--\eqref{eq:gains}. From Balanced fidelity up the
  propagation matrix carries the seeded phase-mismatch term,
  Eq.~\eqref{eq:Mdeltak}; High fidelity adds the refractive correction and the
  16-segment absorption profile; Ultra runs the 64-segment dynamic
  depletion path with Gaussian overlap (Sec.~\ref{sec:ultraprop}).
  \item Apply the Manley--Rowe cap, Eq.~\eqref{eq:manleyrowe}.
  \item Squeezing $S_{\text{dB}}(\delta)$: Eq.~\eqref{eq:Seta} on the
  legacy/Balanced/Fine paths; the hardened balanced-arm noise
  Eq.~\eqref{eq:hardenedS} on Ultra.
\end{enumerate}
The returned dictionary carries $G_s$, $G_c$, $S_{\text{dB}}$ versus probe
detuning, plus derived diagnostics: $N$, $\sigma_v$, $\Omega_{\text{pump}}$,
$\Omega_{\text{seed}}$, the small-signal peak and the depletion cap, the
first-principles \code{coupling\_norm} $\varrho_{\text{norm}}$, the
\code{line\_strength\_residual} $\ell_s$ and their product
(\code{effective\_line\_strength}), the active phase-detail mode, $\Delta k_z$
(operating and vacuum), segment/overlap/pump-budget diagnostics, the Zeeman
consistency status, and --- on Ultra --- the \code{hardened\_noise} block
($\kappa$, pump OD, per-arm linear ODs). The TPD slider in the app reads a
single operating point off this curve by interpolation
(\code{operating\_point}) --- instant, no re-solve --- which is why the $\delta$
knob recomputes nothing.

\paragraph{Anchor.} The direct \code{compute\_spectrum} path remains
regression-anchored (\code{tests/test\_regression.py}, baseline
\code{tests/baseline\_focused.npz}; re-captured after the coupling
normalization, the collisional dissipator, and the IDS-formula correction). At
the Sim \emph{et al.}\ reference setting ($\Delta=0.9\,$GHz, $T=121\,^\circ$C,
$600\,$mW pump, $8\,\mu$W seed, $\ell_s=0.74$, $5.5\,\%$ loss,
$\eta=0.8694$), the pinned legacy focused scan reaches
$S_{\text{dB}}\approx-8.69\,$dB at $G_s\approx32.1$, against the
$-8.84\,$dB detection floor of Eq.~\eqref{eq:floor}. The same residual
$\ell_s=0.74$ makes the \emph{Ultra} path reproduce the paper's measured
operating point ($G_s\approx14$, $-7.8\,$dB) --- that is the calibration
statement; the legacy path, which omits the propagation refinements, runs
hotter. Regression tests hold the seeded gain/squeezing arrays to the frozen
baseline bit-for-bit, and the app presets select the phase-detail mode per
fidelity (Fast/Balanced $\to$ \code{balanced}, High $\to$ \code{fine},
Ultra $\to$ \code{ultra}).

% =====================================================================
\section{Spontaneous FWM: the biphoton mode}
% =====================================================================

The second mode (\code{GenericFWMSolver}) drops the seed entirely and estimates the
\emph{spontaneous} biphoton source: pairs generated from vacuum in a cascade or
diamond level topology. It now carries two source models
(\code{biphoton\_model} toggle): the default \textbf{Predictive} path solves the
Doppler-averaged biphoton amplitude from first principles --- waveform shape,
temporal width, source bandwidth and their parameter dependences are
\emph{computed} --- while \textbf{Calibrated} is the legacy reference-injected
estimate kept for comparison. In both, the absolute pair rate stays
reference-anchored with physical scaling: the lumped model does not pin the
absolute collection coefficient, for the same reason the squeezing mode keeps
the $\ell_s$ residual (Sec.~\ref{sec:coupling}).

\subsection{Topologies and fields}

A topology is a set of \code{LevelSpec} levels and \code{FieldSpec} fields (pump,
coupling, signal, idler), each with a wavelength, propagation direction, and
phase sign. The driven-field Rabi frequencies are scaled from per-topology
reference anchors as $\sqrt{P/P_{\text{ref}}}$
(\code{beam.anchored\_rabi\_mhz}), so the power knobs move the physics without
pretending to an absolute beam-geometry calibration. Three topologies are
provided:
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{$^{87}$Rb telecom cascade} $5S_{1/2}\!\to\!5P_{3/2}\!\to\!4D_{5/2}$
  (780/1529 nm), calibrated to $\sim$38\,000 cps/mW, $g^{(2)}\!\approx\!44$,
  OD$\,\approx\,$112, bandwidth $\approx300\,$MHz.
  \item \textbf{$^{133}$Cs ``BTW'' cascade}, two channels (852--917 nm and
  852--795 nm) that differ mainly in their temporal width.
  \item \textbf{Generic diamond} four-level template with user wavelengths (no
  validated anchor).
\end{itemize}

\subsection{Energy and vector momentum conservation}

Energy conservation across the four fields is a bookkeeping check
(\code{energy\_mismatch\_hz}):
\begin{equation}
\Delta\nu = \nu_{\text{pump}}+\nu_{\text{coupling}}-\nu_{\text{signal}}-\nu_{\text{idler}}\ \approx\ 0.
\end{equation}
Momentum conservation (phase matching) is evaluated as a vector mismatch
(\code{phase\_mismatch\_vector}). The longitudinal component remains calibrated
to the reference geometry,
\begin{equation}
\Delta k_z^{\text{rel}}
=\Delta k_z^{\text{abs}}-\Delta k_{\text{ref}},
\qquad
\Delta k_z^{\text{abs}}
=\sum_{\text{fields}}s_i\,k_i\cos\theta_i,
\qquad k_i=\frac{2\pi}{\lambda_i},
\end{equation}
with phase signs $s_i$ and collection angles $\theta_i$. The reference
$\Delta k_{\text{ref}}$ is applied only to the longitudinal component. The
transverse component is absolute:
\begin{equation}
\Delta k_x=\sum_{\text{fields}}s_i\,k_i\sin\theta_i\,\sigma_i,
\qquad
\sigma_i\in\{-1,0,+1\}.
\end{equation}
Pump and coupling fields use $\sigma_i=0$ in the current biphoton collection
model. Signal and idler use the selected collection side. Thus a longitudinally
calibrated but transversely inconsistent setting, such as forcing the
$^{87}$Rb telecom source to $1.5^\circ/1.5^\circ$ on the same side, carries a
large transverse mismatch and is suppressed by the sinc weight.
The finite interaction length $L$ turns this into a collection efficiency through
the standard phase-matching sinc (\code{phase\_matching\_weight}):
\begin{equation}
\Delta k_{\text{vec}}=
\sqrt{\left(\Delta k_z^{\text{rel}}\right)^2+\Delta k_x^2},
\qquad
\Phi_{\text{vec}}=\sinc^2\!\Big(\frac{\Delta k_{\text{vec}}\,L}{2}\Big)
=\left[\frac{\sin(\Delta k_{\text{vec}}\,L/2)}
{\Delta k_{\text{vec}}\,L/2}\right]^2.
\label{eq:sinc}
\end{equation}
Scanning the idler angle traces out the one-dimensional acceptance lobe
$\Phi_{\text{vec}}(\theta_{\text{idler}})$. In Fine phase-detail mode, the same
vectorized geometry is evaluated on a signal--idler grid
(\code{phase\_mismatch\_vector\_grid}), producing a two-dimensional acceptance map
$\Phi_{\text{vec}}(\theta_s,\theta_i)$ without touching the velocity-class source
calculation.

\subsection{Predictive first-principles waveform}
\label{sec:predictive}

The predictive model (\code{\_predictive\_waveform}) evaluates the biphoton
temporal amplitude as a \emph{coherent sum over Maxwell velocity classes} --- the
hot-vapour analogue of the joint spectral amplitude --- following Kim
\emph{et al.}\ [\emph{Quantum Sci.\ Technol.}\ \textbf{9}, 045006 (2024),
Eq.~(2)], Chen \emph{et al.}\ [\emph{Phys.\ Rev.\ Research} \textbf{4}, 023132
(2022), Eqs.~(3--5)] and Du, Wen, Rubin [\emph{J.\ Opt.\ Soc.\ Am.\ B}
\textbf{25}, C98 (2008)]. Each velocity class contributes the two-photon
amplitude
\begin{equation}
\kappa(v)=w(v)\,
\frac{\Omega_p\,\Omega_c}
{4 f_1(v)\,f_2(v)+\Omega_c^2}\;\propto\;\mathrm{OD},
\qquad
\begin{aligned}
f_1(v)&=\tfrac{\Gamma_i}{2}+\ii\big(\Delta_p-k_p v\big),\\
f_2(v)&=\tfrac{\Gamma_e}{2}+\ii\big(\Delta_{2\gamma}-k_{\text{res}} v\big),
\end{aligned}
\label{eq:kappa}
\end{equation}
with $\Gamma_i,\Gamma_e$ the intermediate/upper natural linewidths,
$\Delta_{2\gamma}=\Delta_p+\Delta_c$ the two-photon detuning, and
$k_{\text{res}}=k_p+k_c$ (signed by propagation directions) the residual
two-photon Doppler wavevector. Three points carry the physics:
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{The $\Omega_c^2$ Autler--Townes term} in the denominator is kept
  --- no weak-coupling approximation --- so a strong coupling laser splits the
  two-photon resonance and reshapes (shortens) the waveform, as in the
  references.
  \item \textbf{The nonlinear response is $\propto$ OD}: $\kappa\propto\alpha$,
  so the pair \emph{rate} scales as $\mathrm{OD}^2$ before collection factors.
  \item \textbf{Two-photon (ground) dephasing} $\gamma_g=0.02\,\Gamma_i$
  (\code{GROUND\_DEPHASING\_FRAC}; Chen \emph{et al.}\ fit
  $\gamma\approx0.02$--$0.03\,\Gamma$) enters the longitudinal response below.
\end{itemize}
The time-domain nonlinear response is the coherent velocity sum with the
single-photon Doppler phase and the \emph{natural} decay envelope of the
idler-emitting intermediate state --- no injected lifetime:
\begin{equation}
\tilde\kappa(\tau)=\Big[\sum_v \kappa(v)\,e^{\,\ii k_p v\,\tau}\Big]\,
e^{-\Gamma_i\tau/2},
\qquad
\psi(\tau)=\tilde\kappa(\tau)\,\sqrt{\Phi_{\text{vec}}},
\label{eq:predpsi}
\end{equation}
where the velocity sum is a Doppler-dephasing integral (classes accumulate
phase $k_p v\tau$ and wash out) whose interference sets the beat-then-decay
biphoton temporal-waveform (BTW) shape; the wavelength-dependent width
\emph{ordering} of the two Cs BTW channels (917 narrower than 795) emerges from
this sum rather than being injected. The source bandwidth is measured from the
waveform's own spectrum, and a regime diagnostic compares the Du group-delay
time $\tau_g=(2\gamma_g/\Omega_c^2)\,\mathrm{OD}\,\Gamma_e$ against the Rabi
time $2\pi/\Omega_c$ (``group-delay'' vs ``damped-Rabi'' regime). Chen's
ultimate spectral-brightness ceiling, $\approx(\pi/2)\times10^{6}$
pairs\,s$^{-1}$\,MHz$^{-1}$, is reported alongside.

\paragraph{Optional OD reshaping (off by default).} At high optical depth the
pair propagates through its own EIT/slow-light medium: in the conjugate
(frequency) domain the joint amplitude is multiplied by the longitudinal
function $\Phi(\delta)=\sinc\varrho\;e^{\ii\varrho}$ [Du Eq.~(14)], with
$\varrho(\delta)$ the OD-weighted complex EIT phase [Chen Eq.~(5)],
\begin{equation}
\varrho(\delta)=\frac{\mathrm{OD}\,\Gamma_e}{2}
\sum_v w(v)\,
\frac{\delta+\ii\gamma_g}
{\Omega_c^2-4\big(\delta+\ii\gamma_g\big)
 \big(\delta+\Delta_{2\gamma}-k_{\text{res}}v+\ii\Gamma_e/2\big)},
\label{eq:eitphase}
\end{equation}
implemented as an FFT convolution of $\tilde\kappa(\tau)$
(\code{\_apply\_longitudinal\_response}), with $\varrho$ clipped to the
dilute-regime range. At $\mathrm{OD}\to0$, $\Phi\to1$ and
$\psi\to\tilde\kappa$ exactly. The path is a diagnostic opt-in
(\code{biphoton\_od\_reshaping}) because the lumped model overestimates the
group-delay broadening at high OD and would smear the validated narrow telecom
BTW.

\subsection{Velocity-grid auto-refinement}
\label{sec:vrefine}

The velocity sum in Eq.~\eqref{eq:predpsi} is a discretized Fourier integral of
the \emph{narrow} source resonance $|\kappa(v)|$, whose velocity-space width is
of order $\Gamma/k$ --- a few m/s, far narrower than the Doppler width
$\sigma_v$ the navigate-only \code{biphoton\_velocity\_step} knob is sized for.
A step coarser than that resonance \emph{aliases} the sum: before the fix, the
reported telecom BTW width swung by a factor $\sim20$, non-monotonically, as
the user step varied over $1$--$12$\,m/s --- numerics masquerading as physics.
The predictive path therefore treats the user step only as an upper bound
(\code{\_converged\_predictive\_waveform}):
\begin{enumerate}[leftmargin=2.2em]
  \item Measure the actual resonance FWHM of $|\kappa(v)|$ on a fine probe grid
  (so Autler--Townes and detuning broadening are captured), and start from
  $\dd v=\mathrm{FWHM}/16$ (\code{PRED\_V\_OVERSAMPLE}).
  \item Halve the step until the $|\psi|^2$ FWHM is stable to $3\,\%$
  (\code{PRED\_V\_FWHM\_TOL}), up to 10 halvings.
  \item If the $40\,000$-point cap is reached first, flag the width as
  \emph{unconverged} (qualitative only) in the readout.
\end{enumerate}
With this in place the $^{87}$Rb telecom predictive width converges to
$\approx0.50$\,ns against the measured $0.56(4)$\,ns, and the Cs channels land
within $\sim30\,\%$ --- the remaining gap is the deferred per-source
absolute-width calibration (\code{docs/checklist.json}).

\subsection{Absolute rate and optical depth: anchored, physically scaled}

The absolute pair rate is reference-anchored with physical scaling
(\code{compute\_biphoton}):
\begin{equation}
R_{\text{pair}}
=R_{\text{ref}}\;P_{\text{pump}}[\mathrm{mW}]\;
\sqrt{s_c}\;\;\Phi_{\text{vec}}\;
\sqrt{N/N_{\text{ref}}},
\label{eq:pairrate}
\end{equation}
with $R_{\text{ref}}$ the per-topology anchor (e.g.\ $38\,000$\,cps/mW for the
telecom cascade), $s_c$ the relative coupling drive, $\Phi_{\text{vec}}$ the
vector phase-matching weight of Eq.~\eqref{eq:sinc}, and $N(T)$ the isotope
density. The optical depth that drives Eqs.~\eqref{eq:kappa} and
\eqref{eq:eitphase} is likewise the \emph{measured} reference OD scaled by
density and cell length, $\mathrm{OD}=\mathrm{OD}_{\text{ref}}\,
(N/N_{\text{ref}})(L/L_{\text{ref}})$ --- OD is an experimentally reported
quantity in these sources, like the cell temperature. Where a topology carries
no reference OD, a physical leg OD is computed from the natural linewidth via
the AutoOD-validated $\Gamma\to|d|^2$ route,
$\alpha_0=2Nk|d|^2/(\varepsilon_0\hbar\Gamma)$
(\code{species.reduced\_dipole\_sq}).

\subsection{Calibrated (legacy) source model}

The Calibrated model keeps the pre-2026-06-19 estimate for comparison: the
velocity-class source is the weak-drive product of the two cascade resonances,
\begin{equation}
A(v)=w(v)\,\frac{\Omega_{p}\,\Omega_{c}}
{\big[\tfrac{\gamma_\ell}{2}+\ii(\Delta_{p}-k_{p}v)\big]\,
 \big[\tfrac{\gamma_u}{2}+\ii(\Delta_{2\gamma}-k_{\text{res}}v)\big]}\,
 \sqrt{\Phi_{\text{vec}}},
\label{eq:sourcev}
\end{equation}
summed with the idler-arm phase $e^{\ii k_{\text{idler}}v\tau}$ and multiplied
by an \emph{injected} exponential lifetime envelope
$e^{-\tau/\tau_{\text{decay}}}$ (the topology's reference emission decay); the
source bandwidth is the reference value and the $g^{(2)}$ peak is forced to the
reference anchor (below). Because the width is set by the injected lifetime
rather than the velocity-sum dephasing, this model is insensitive to the
velocity-step aliasing of Sec.~\ref{sec:vrefine}.

\subsection{Correlation readout and calibration}

The detector model (\code{observables.biphoton\_stats}) folds in finite filter
transmission $\eta_f=\min(1,\,\text{BW}_{\text{filter}}/\text{BW}_{\text{source}})$,
detector efficiencies $\eta_{s,i}$, dark counts, a coincidence window, and timing
jitter (Gaussian smoothing of $|\psi|^2$). It then forms singles, true and
accidental coincidences, and the normalized cross-correlation
\begin{equation}
g^{(2)}_{SI}(\tau)=1+\text{CAR}\cdot\frac{|\psi(\tau)|^2}{\max|\psi|^2},
\qquad
\text{CAR}=\frac{\text{coincidence}}{\text{accidental}}.
\end{equation}
Heralding efficiencies and the Cauchy--Schwarz ratio $R=[g^{(2)}_{\text{peak}}]^2/4$
follow. The two source models differ in one honest detail: in the
\emph{Calibrated} model, where a paper anchor exists (e.g.\ the $^{87}$Rb
$g^{(2)}\!\approx\!44$), an explicit added-accidental term calibrates the peak
to the reference value --- the absolute noise floor is \emph{calibrated}, not
derived. The \emph{Predictive} model passes no $g^{(2)}$ target: its peak
follows from the computed waveform and the physical accidentals alone. The
in-app validation table marks each anchor accordingly --- calibrated entries as
``by construction'', predictive entries as ``predicted (approx)'' --- and checks
rate/OD/width anchors, energy bookkeeping, phase matching, and the velocity
convergence flag.

% =====================================================================
\section{Numerical architecture (brief)}
% =====================================================================

The physics dictates the architecture, so a short note suffices:
\begin{itemize}[leftmargin=1.6em]
  \item \textbf{One template, shifted.} $\Lcal_0(\Delta_\eff)=\Lcal_0(0)-\Delta_\eff
  S_v$ (Eq.~\eqref{eq:deffshift}) lets every velocity share one Liouvillian and be
  assembled by cheap matrix shifts.
  \item \textbf{$\Delta$-independent, per-$T$ table.} $\rhobar(\delta,\Delta_\eff)$
  is computed on the needed grid; $\Delta\to$ sampled columns or interpolation
  coordinates, $T\to$ Maxwell weights \emph{plus} the collisional dissipator
  (Sec.~\ref{sec:collisional}). The $(\Delta,T)$ map therefore builds one
  exact-union master grid per temperature and reuses it across all $\Delta$.
  \item \textbf{Compiled OBE kernels.} \code{gabes/kernels.py} provides optional
  Numba kernels for the hot solve grids. FWM uses
  \code{floquet\_chi\_grid}; the absorption/Rydberg family uses the affine-scan
  kernel. The NumPy paths remain the reference fallback when Numba is unavailable.
  \item \textbf{Closed-form $2\times2$.} The transfer matrix exponential is
  analytic (Appendix~\ref{app:expm}), avoiding a per-point \code{expm}; its
  exponent arguments are overflow-clamped so runaway linear gain degrades to
  huge-but-finite values instead of NaN-poisoning the curve (Sec.~\ref{sec:mb}).
  \item \textbf{Thread hygiene.} The small dense systems are kept out of
  oversubscribed BLAS threading with \code{core.blas\_single\_thread}; compiled
  kernels parallelize at the grid level, and \code{full\_spectrum} computes the
  two Raman branches independently.
  \item \textbf{Headless readout.} Every scheme's \code{observables} accepts
  \code{include\_figures=False}, returning metrics/tables without building
  Matplotlib figures --- the batch scanners and reports use this path (figure
  construction, not the OBE solve, dominates a warmed interactive call).
\end{itemize}

% =====================================================================
\section{Validation, scope, and what is \emph{not} modelled}
% =====================================================================

\paragraph{Validated / first-principles.} The direct seeded
$G_s$/$S_{\text{dB}}$ pipeline is regression-locked bit-for-bit, and at Ultra
fidelity reproduces the Sim \emph{et al.}\ measured operating point
($G_s\approx14$, $-7.8\,$dB) with the single residual $\ell_s=0.74$. The
optional Numba kernels are pinned to the NumPy reference paths, and the
app-facing phase-detail modes add propagation physics without changing the
atomic OBE definition. The genuine $^{85}$Rb $D_1$ hyperfine CG factors
$(10,35,35,28)/81$ drive the coherent coupling, the polarization readout, and
the spontaneous-emission branching; the macroscopic scale chain --- CRC density,
$p_F/[2(2I+1)]$ normalization, self-broadening $\beta$, and the hardened
arm/pump absorption channels --- is tied to the AutoOD-validated absorption path
($<0.1\,\%$ against measured cell transmission). The ideal
intensity-difference noise is the conserved-difference result
Eq.~\eqref{eq:Sideal}, and the collisional dissipator gives the observed
interior temperature optimum.

\paragraph{Lumped / calibrated residuals.} What still carries a knob: the
dimensionless line-strength residual $\ell_s=0.74$ ($m$-sublevel participation /
beam geometry, not yet derived); the pump-scatter leakage fraction
$\kappa=0.1$ of the hardened Ultra noise (PBS extinction / spatial filtering,
awaiting measurement); the biphoton absolute pair rate and (in the Calibrated
model) the $g^{(2)}$ noise floor, which are reference-anchored; and the
predictive biphoton \emph{absolute} widths, which are approximate
(telecom $\approx0.50$ vs $0.56(4)$\,ns; Cs within $\sim30\,\%$) pending the
per-source calibration deferred in \code{docs/checklist.json}.

\paragraph{Ultra status and remaining future physics.}
\begin{itemize}[leftmargin=1.6em]
  \item Ultra includes single-pass refractive phase matching, Gaussian
  beam-overlap weighting, 64-segment dynamic pump-budget depletion, in-cell
  beam-splitter loss before detector efficiency, and --- intrinsic since v5 ---
  the hardened excess-noise model (arm-resolved linear absorption, OD-anchored
  pump scatter). \code{excess\_noise\_model=False} reproduces pre-hardening
  Ultra for A/B.
  \item Remaining zero-default technical noise channels (seed excess noise, EOM
  residual multimode noise) await calibrated measurements.
  \item Full Zeeman ($m_F$) substructure is available as a 24-level $^{85}$Rb
  $D_1$ diagnostic manifold whose CG sums reproduce the lumped FWM strengths
  (consistency $\equiv1.0$, reported but not multiplied in). The full 24-level
  Floquet scan ($M=576$ vs $M=16$) remains future work.
  \item Optical-pumping depletion of the $F=3$ ground manifold when the pump
  parks near its resonances --- the leading candidate for the experiment's
  $+0.9\,$GHz preference over the model's shallow $-2.0\,$GHz optimum --- is not
  representable in the lumped steady-state model (Sec.~\ref{sec:hardened}).
  \item Full quantum-Langevin propagation of loss/noise (beyond the
  beamsplitter treatment) remains future work; the twin-beam photon statistics
  stay ideal-parametric estimates.
\end{itemize}

% =====================================================================
\appendix
\section{Symbol glossary}
% =====================================================================

\begin{center}
\begin{tabular}{ll}
\toprule
symbol & meaning \\
\midrule
$\Delta$ & one-photon (pump) detuning from $F=2\!\to\!F'=3$ \\
$\delta$ & two-photon (Raman) detuning \\
$\Delta_\eff=\Delta-kv$ & effective one-photon detuning of a velocity class \\
$\Omega$ & Rabi frequency (drive strength) \\
$\nu_{\HF}=3.0357\,$GHz & ground hyperfine splitting \\
$\Omega_{\text{beat}}=\Omega_{\HF}\mp\delta$ & ground-coherence beat (Floquet) frequency \\
$\Gamma/2\pi=5.746\,$MHz & natural linewidth \\
$\gamma_{gg}(T,N)$ & ground Raman dephasing: $2\pi\!\cdot\!100\,$kHz floor $+\,N\sigma\bar v_{\text{rel}}$, Eq.~\eqref{eq:gammagg} \\
$\gamma_{\text{opt}}=\beta N/2$ & collisional optical dephasing (self-broadening), Eq.~\eqref{eq:gammaopt} \\
$C_F^2$ & hyperfine Clebsch--Gordan line strength \\
$\rhobar_{xy}$ & dimensionless susceptibility (leg $x$, field $y$) \\
$\chi_{xy}$ & physical susceptibility, Eq.~\eqref{eq:chiphys} \\
$N$ & atomic number density (pure-$^{85}$Rb CRC, Eq.~\eqref{eq:crc}) \\
$T$ (matrix) & $2\times2$ field transfer matrix $\exp(ML)$ \\
$G_s,G_c$ & seed / conjugate power gain \\
$S_{\text{ideal}}=(G_s-G_c)^2/(G_s+G_c)$ & ideal twin-beam noise, Eq.~\eqref{eq:Sideal} \\
$S,S_{\text{dB}}$ & intensity-difference squeezing (linear / dB) \\
$\eta=\text{QE}(1-\text{loss})$ & detection efficiency ($0.8694$ at reference) \\
$\varrho_{\text{norm}}=p_F/[2(2I+1)]$ & first-principles coupling norm ($7/144$ on branch $-$) \\
$\ell_s$ & residual line-strength calibration (\code{LINE\_STRENGTH\_FACTOR}; app $0.74$) \\
$\eta_{s,c}^{\text{arm}}$ & hardened per-arm efficiencies, Eq.~\eqref{eq:armloss} \\
$\kappa$ & pump-scatter leakage fraction (hardened Ultra, default $0.1$) \\
$\Delta k_z$ & seeded longitudinal phase mismatch \\
$\Delta k_z^{\text{abs}},\Delta k_z^{\text{rel}}$ & biphoton longitudinal absolute / calibrated relative mismatch \\
$\Delta k_x,\Delta k_{\text{vec}}$ & biphoton transverse and vector phase mismatch \\
$\Phi_{\text{vec}}=\sinc^2(\Delta k_{\text{vec}} L/2)$ & biphoton vector phase-matching weight \\
$\psi(\tau)$ & biphoton temporal waveform \\
\bottomrule
\end{tabular}
\end{center}

% =====================================================================
\section{Closed-form \boldmath$2\times2$ matrix exponential}
\label{app:expm}
% =====================================================================

Write $A=s\,\mathbb{1}+q$ with $s=\tfrac12\Tr A$ so that $q$ is traceless. By the
Cayley--Hamilton theorem a traceless $2\times2$ matrix satisfies
$q^2=-\det(q)\,\mathbb{1}=c^2\,\mathbb{1}$ with $c=\sqrt{-\det q}$. Hence
\begin{equation}
e^{qL}=\cosh(cL)\,\mathbb{1}+\frac{\sinh(cL)}{c}\,q,
\qquad
e^{AL}=e^{sL}\Big[\cosh(cL)\,\mathbb{1}+\frac{\sinh(cL)}{c}\,q\Big].
\end{equation}
This is evaluated entrywise and batched over all detuning points in
\code{core.matrix\_exp\_2x2}; the $c\to0$ limit (where $\sinh(cL)/c\to L$) is
handled explicitly to avoid the removable singularity. For the propagation matrix
$M$ of Eq.~\eqref{eq:Mmatrix} this gives the transfer matrix $T=e^{ML}$ exactly,
with no series truncation. The real parts of $cL$ and $sL$ are clipped at
$\pm350$ before $\cosh/\sinh/\exp$ (two factors multiply; float64 overflows near
$709$): an unphysical runaway linear gain then saturates finite and is capped by
the Manley--Rowe stage, instead of overflowing to \code{inf} and emerging as NaN
from $\infty/(1+\infty)$ in the saturation formula.

% =====================================================================
\section{File / function map}
% =====================================================================

\begin{center}\footnotesize
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lll}
\toprule
physics & function & file \\
\midrule
Rabi from power & \code{rabi\_freq} & \code{constants.py} \\
density $N(T)$ (seeded, pure-$^{85}$Rb CRC) & \code{number\_density} & \code{hyperfine.py} \\
density (other schemes / biphoton) & \code{rb85\_density}, \code{number\_density} & \code{atoms.py}, \code{species.py} \\
CG line strengths & \code{CF2}, \code{line\_strength}, \code{cf2} & \code{hyperfine.py}, \code{species.py} \\
coupling norm $p_F/[2(2I+1)]$ & \code{physical\_coupling\_norm} & \code{schemes/fwm.py} \\
level model + Lindblad & \code{AtomModel}, \code{double\_lambda\_rb85} & \code{atoms.py} \\
collisional atom (per $T$) & \code{collisional\_atom} & \code{schemes/fwm.py} \\
ground Raman rate $\gamma_{gg}(T,N)$ & \code{ground\_coherence\_dephasing} & \code{constants.py} \\
self-broadening $\Gamma+\beta N$ & \code{self\_broadened\_gamma} & \code{hyperfine.py} \\
commutator super-op & \code{comm\_super} & \code{core.py} \\
Liouvillian & \code{build\_liouvillian} & \code{core.py} \\
Hamiltonians & \code{static\_hamiltonian\_at\_Deff\_zero}, \code{sideband\_hamiltonian} & \code{schemes/fwm.py} \\
Floquet steady state & \code{floquet\_solve} & \code{core.py} \\
$\rhobar$ matrix table & \code{chi\_matrix\_table}, \code{floquet\_chi\_grid} & \code{schemes/fwm.py}, \code{kernels.py} \\
velocity grid / $\Delta_\eff$ & \code{velocity\_grid}, \code{build\_Delta\_eff\_axis} & \code{doppler.py} \\
Doppler average & \code{doppler\_average} & \code{doppler.py} \\
Maxwell--Bloch gain / segmented propagation & \code{gain\_from\_chi} & \code{observables.py} \\
$2\times2$ matrix exp (overflow-clamped) & \code{matrix\_exp\_2x2} & \code{core.py} \\
pump-depletion cap & \code{pump\_depletion\_saturation} & \code{observables.py} \\
ideal twin-beam noise & \code{ideal\_twin\_beam\_noise} & \code{observables.py} \\
squeezing (symmetric $\eta$) & \code{intensity\_difference\_squeezing\_dB} & \code{observables.py} \\
balanced two-arm noise & \code{balanced\_twin\_beam\_noise} & \code{observables.py} \\
in-cell loss + technical noise & \code{segmented\_loss\_noise\_squeezing\_dB} & \code{observables.py} \\
arm linear OD (hardened) & \code{\_arm\_linear\_od} & \code{schemes/fwm.py} \\
pump-scatter noise (hardened) & \code{\_pump\_scatter\_noise} & \code{schemes/fwm.py} \\
Ultra segmented propagation & \code{\_ultra\_segmented\_gain} & \code{schemes/fwm.py} \\
twin-beam statistics & \code{coincidence\_stats} & \code{observables.py} \\
full seeded pipeline & \code{compute\_spectrum} & \code{schemes/fwm.py} \\
biphoton source & \code{GenericFWMSolver.compute\_biphoton} & \code{schemes/fwm.py} \\
predictive biphoton waveform & \code{\_predictive\_waveform}, \code{\_converged\_predictive\_waveform} & \code{schemes/fwm.py} \\
seeded $\Delta k_z$ & \code{seeded\_phase\_mismatch\_z} & \code{schemes/fwm.py} \\
phase matching & \code{phase\_mismatch\_vector}, \code{phase\_mismatch\_vector\_grid}, \code{phase\_matching\_weight} & \code{schemes/fwm.py} \\
biphoton readout & \code{biphoton\_stats} & \code{observables.py} \\
\bottomrule
\end{tabular}
\end{center}

\vfill
\begin{center}\small
Reference: G.\ Sim, H.\ Kim, H.\ S.\ Moon, \emph{Sci.\ Rep.}\ \textbf{15}, 7727 (2025);
Rb data from D.\ A.\ Steck, ``Rubidium 85 D Line Data.''
\end{center}

\end{document}
